---
title: "Covid91 vaccine study Final2024"
author: "Jasmin Quintanilla"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The covid921 vaccine is not effective in males 

$H_A$: The covid921 vaccine is effective in males 


### Methods

### Descriptive results

### Graphical results 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

This chart indicates that individuals who received the placebo have a higher percentage of contracting COVID-19 compared to those who received the vaccine.

#### Numerical results 

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This table demonstrates that the placebo group has a higher incidence of contracting COVID-19, with a 70% infection rate. In contrast, the vaccine group shows a significantly lower infection rate of 30%, indicating that the vaccine is a more effective alternative for reducing the risk of COVID-19 in males.


#### Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi and fisher have the same p-value. The p-value is 0.000000000000002 which means its lower than 0.05.

### Conclusion

The chart shows that individuals who received the placebo have a higher chance of getting COVID-19 compared to those who received the vaccine. The table indicates that 70% of the placebo group got infected, while only 30% of the vaccine group did. This suggests that the vaccine is more effective at reducing the risk of COVID-19 in males. Both the chi-square test and Fisher's exact test have the same p-value of 0.000000000000002, which is much lower than the 0.05 threshold. This strong statistical evidence supports the conclusion that the COVID-19 vaccine is effective in males, as the placebo group has a higher infection rate.

## Females

$H_0$: The covid921 vaccine is not effective in females

$H_A$: The covid921 vaccine is effective in females 

### Methods



### Descriptive results

### Graphical results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This chart indicates that individuals who received the placebo have a higher percentage of contracting COVID-19 compared to those who received the vaccine.

#### Numerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```

This table indicates that the vaccine reduces the risk of contracting COVID-19 to 40% in females, whereas the placebo group shows a 60% risk of infection. Based on this data, it is advisable to choose the vaccine, as it significantly lowers the probability of contracting COVID-19.

#### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi and fisher have the same p-value. The p-value is 0.000000000000002 which means its lower than 0.05.

### Conclusion 

This chart shows that individuals who received the placebo have a higher chance of getting COVID-19 compared to those who received the vaccine. The data indicates that the vaccine reduces the risk of infection to 40% in females, while the placebo group has a 60% risk. Therefore, it is better to choose the vaccine as it significantly lowers the chance of contracting COVID-19. Both the chi-square test and Fisher's exact test give the same p-value of 0.000000000000002, which is much lower than 0.05. This strong statistical evidence supports the conclusion that the COVID-19 vaccine is effective in females, as the placebo group has a higher infection rate.

## LGBTQ

$H_0$: The covid921 vaccine is not effective in LGBTQ Community

$H_A$: The covid921 vaccine is effective in LGBTQ Community

### Methods

### Descriptive results

### Graphical results

```{r}
LGBTQ=filter(FinalData, LGBTQ=="gay")
```



```{r}
barchartGC(~treatment +infected ,data=LGBTQ)
barchartGC(~treatment +infected ,data=LGBTQ,type="percent")

```

The graph indicates that individuals who are part of the LGBTQ community have similar outcomes when taking either the placebo or the vaccine, showing no significant difference in COVID-19 infection rates. This suggests that neither the placebo nor the vaccine has a notably different effect on this group.

#### Numerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
When looking at this table it shows that the placebo has a 15% of getting covid19 well as the vaccine would be at an 85% chance of getting covid19 for the LGBTQ Community. In this case we would choose the placebo instead of the vaccine. 


#### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi and fisher have the same p-value. The p-value is 0.00000006 which means its higher than 0.05.


### Conclusion

The graph shows that individuals in the LGBTQ community have similar COVID-19 infection rates whether they receive the placebo or the vaccine. This means neither treatment has a notably different effect. The table indicates that the placebo group has a 15% infection rate, while the vaccine group has an 85% infection rate. Based on this data, the placebo seems to be a better option. Both the chi-square test and Fisher's exact test have the same p-value of 0.00000006, which is much lower than 0.05. This suggests that the null hypothesis—that there is no difference between the placebo and the vaccine in this group—can be rejected. Therefore, the vaccine is considered ineffective for the LGBTQ community according to these findings.

## Druggies

$H_0$: The covid921 vaccine  is not effective in druggies

$H_A$: The covid921 vaccine is effective in druggies

### Methods

### Descriptive results

### Graphical results 

```{r}
druggies=filter(FinalData,FinalData$DrugUser=="yes")
barchartGC(~treatment + infected,data=druggies)
barchartGC(~infected + treatment,data=druggies,type="percent")

```

This graph demonstrates that individuals who use drugs and received the vaccine have a higher incidence of COVID-19 infection, indicating that the placebo is a safer option for this group.

#### Numerical results 

```{r}
table1 <- xtabs(~treatment + infected,data=druggies)
rowPerc(table1)
colPerc(table1)
```

The table indicates that individuals who use drugs have an 88% chance of contracting COVID-19 when given the vaccine, compared to an 11% chance when given the placebo. 

#### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi-square test and Fisher's exact test yield the same p-value of 0.0000000002, which is significantly lower than the 0.05 threshold.

### Concluion

This graph demonstrates that individuals who use drugs and received the vaccine have a higher incidence of COVID-19 infection, indicating that the placebo is a safer option for this group. The table shows that these individuals have an 88% chance of contracting COVID-19 when given the vaccine, compared to an 11% chance when given the placebo. Both the chi-square test and Fisher's exact test yield the same p-value of 0.0000000002, which is significantly lower than the 0.05 threshold. Therefore, we reject the hypothesis that the vaccine is effective for individuals who use drugs; the placebo proves to be more effective in this case.


