---
title: "Covid91 vaccine study Final2024"
author: "Jasmin Quintanilla"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The covid921 vaccine is not effective in males

$H_A$: The covid921 vaccine is effective in males 


### Methods

### Descriptive results

### Graphical results 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

This chart shows that the people who took the

#### Numerical results 

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This table demonstrates that the placebo group has a higher incidence of contracting COVID-19, with a 70% infection rate. In contrast, the vaccine group shows a significantly lower infection rate of 30%, indicating that the vaccine is a more effective alternative for reducing the risk of COVID-19 in males.


#### Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi and fisher have the same p-value. The p-value is 0.000000000000002 which means its lower than 0.05.

### Conclusion



## Females

$H_0$: The covid921 vaccine is not effective in females

$H_A$: The covid921 vaccine is effective in females 

### Methods



### Descriptive results

### Graphical results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This graph shows that...

#### Numerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```

This table indicates that the vaccine reduces the risk of contracting COVID-19 to 40% in females, whereas the placebo group shows a 60% risk of infection. Based on this data, it is advisable to choose the vaccine, as it significantly lowers the probability of contracting COVID-19.

#### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi and fisher have the same p-value. The p-value is 0.000000000000002 which means its lower than 0.05.

### Conclusion 



## LGBTQ

$H_0$: The covid921 vaccine is not effective in LGBTQ Community

$H_A$: The covid921 vaccine is effective in LGBTQ Community

### Methods

### Descriptive results

### Graphical results

```{r}
LGBTQ=filter(FinalData, LGBTQ=="gay")
```



```{r}
barchartGC(~treatment +infected ,data=LGBTQ)
barchartGC(~treatment +infected ,data=LGBTQ,type="percent")

```

This graph shows...

#### Numerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
When looking at this table it shows that the placebo has a 15% of getting covid19 well as the vaccine would be at an 85% chance of getting covid19 for the LGBTQ Community. In this case we would choose the placebo instead of the vaccine. 


#### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

0.00000006 

### Conclusion



## Druggies

$H_0$: The covid921 vaccine  is not effective in druggies

$H_A$: The covid921 vaccine is effective in druggies

### Methods

### Descriptive results

### Graphical results 

```{r}
druggies=filter(FinalData,FinalData$DrugUser=="yes")
barchartGC(~treatment + infected,data=druggies)
barchartGC(~infected + treatment,data=druggies,type="percent")

```

This graph shows...



